### 2019-03-02 docs/notebooks/results.Rmd

## Abstract
Convolutional neural networks (CNNs) provide cutting edge tools for computer vision tasks like image classification and scene segmentation. Many mechanisms proposed to improve the performance of neural networks have been given the name "attention". Scientists studying the visual system have traditionally thought of attention as an artifact of the limited resources available for processing sensory inputs. Attention in the brain has been primarily investigated using visual search tasks. These tasks may provide an alternate way of benchmarking existing neural network architectures, comparing their output with human behavior, and measuring how any proposed attentional mechanisms change neural network outputs. We are pursuing these alternate benchmarks for attentional mechanisms in the course of our investigations of continual learning algorithms. Here we adapt a classic form of visual search task so it can be performed by neural networks. We measure the accuracy of two CNN architectures, AlexNet and VGG16, as they report whether a target is present among distractors. Surprisingly, we find that both AlexNet and VGG16 show attention-like effects when performing the visual search task. When the target resembles the distractors, then a network's accuracy drops as the number of distractors increases. This dependency of accuracy on the number of items is often referred to as "set size effects". We do not find that these set size effects can be explained by known factors to affect CNNs, such as edge effects. We then show that these attentional effects persist even when controlling for how discriminable the targets are from the distractors, by using an analysis of visual search tasks based on signal detection theory. Hence these "attentional" effects can be attributed to the decision criteria of the network. In the discussion, we consider how this type of analysis can be extended to stimuli besides visual search criteria.

>>>>

## Abstract
Convolutional neural networks (CNNs) provide cutting edge tools for computer vision tasks like image classification and scene segmentation. Many mechanisms proposed to improve the accuracy of CNNs been given the name "attention", although it is becoming clear that accuracy alone does not explain how CNNs perform in real-world situations. Attention has a long history of being studied by cognitive scientists and neuroscientists in the context of vision. The CNNs developed in the past few years are also increasingly being used as models of the visual system in the brain. Here we study whether models of attention in the brain, and visual search tasks used to investigate attention, provide an alternate way of benchmarking existing CNNs. We are pursuing these alternate benchmarks in the course of our investigations of attention-like mechanisms for continual learning algorithms. We adapt a classic form of visual search task so it can be performed by neural networks, and measure the accuracy of two CNN architectures, AlexNet and VGG16, as they report whether a target is present among distractors. Surprisingly, we find that both AlexNet and VGG16 show attention-like effects when performing the visual search task. When the target resembles the distractors, then a network's accuracy drops as the number of distractors increases. This dependency of accuracy on the number of items is often referred to as "set size effects". We do not find that these set size effects can be explained by known factors to affect CNNs, such as edge effects. We then adapt an analysis of visual search tasks based on signal detection theory to show that set size effects persist even when controlling for how discriminable the targets are from the distractors. This analysis suggests the "attentional" effects can be attributed to the decision criteria used by the network. We consider how this type of analysis can be extended to images besides visual search stimuli. Then we discuss the implications of our results for comparing CNN output with human behavior, and measuring how any proposed attentional mechanisms change neural network outputs.
